\Large \bf Uncovering Bugs in Distributed Storage Systems\\during Testing (not in Production!)

### 2016-01-01 evaluation.tex
Edit MigratingTable sections.

For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:migrating} on eleven buggy versions of the system, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect. The test found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}. The remaining four bugs (denoted by $\diamond$) were not caught with our default test harness in the 100,000 executions (a process that required less than 30 minutes). We believe this is because the inputs and schedules that trigger them are too rare in our distribution. To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs, as shown. This method is useful for regression testing, but one could not use it during development with no prior knowledge about the bugs.

The QueryStreamedBackUpNewStream bug in MigratingTable, that was found using \psharp, stands out because it reflects the type of oversight that tends to occur as designs evolve. This bug is in the implementation of a \emph{streaming read} from the virtual table, which should return a stream of all rows in the table sorted by key. The essential implementation idea is to perform streaming reads from both backend tables and merge the results. According to the \texttt{IChainTable} specification, each row read from a stream may reflect the state of the table at any time between when the stream was started and the row was read. The developers sketched a proof that the merging process would preserve this property as long as the migrator was only copying rows from the old table to the new table. But when support was added for the migrator to delete the rows from the old table after copying, it became possible for the backend streams to see the deletion of a row from the old table but not its insertion into the new table, even though the insertion happens first, and the row would be missed. \psharp managed to discover this bug in a matter of seconds. The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening; although, this was after days of experience analyzing traces and after they had extended the logging of \psharp with additional trace information specific to MigratingTable. \psharp only outputs trace information related to its communicating state machines, but the trace information is easy to extend and was done in all our case studies.

>>>>

For the MigratingTable case study, we evaluated the \psharp test harness of \S\ref{sec:migrating} on eleven buggy versions of the system, including eight \emph{organic} bugs that actually occurred in development and three \emph{notional} bugs (denoted by $*$), which are other code changes that we deemed interesting ways of making the system incorrect. The test found seven of the organic bugs, which are listed first in Table~\ref{tab:testing}. The remaining four bugs (denoted by $\diamond$) were not caught with our default test harness in the 100,000 executions (a process that required less than 30 minutes). We believe this is because the inputs and schedules that trigger them are too rare in our distribution. To confirm this, we wrote a custom test case for each bug with a specific input that triggers it and were able to quickly reproduce the bugs;
% Be a little bit more transparent about the fact that the $\diamond$ rows do not
% reflect results achievable via our main testing approach. ~ Matt 2015-12-31
the table shows the results of these runs.
This method is useful for regression testing, but one could not use it during development with no prior knowledge about the bugs.

The QueryStreamedBackUpNewStream bug in MigratingTable, that was found using \psharp, stands out because it reflects a type of oversight that can easily occur as systems evolve. This bug is in the implementation of a \emph{streaming read} from the virtual table, which should return a stream of all rows in the table sorted by key. The essential implementation idea is to perform streaming reads from both backend tables and merge the results. According to the \texttt{IChainTable} specification, each row read from a stream may reflect the state of the table at any time between when the stream was started and the row was read. The developers sketched a proof that the merging process would preserve this property as long as the migrator was only copying rows from the old table to the new table. But when support was added for the migrator to delete the rows from the old table after copying, it became possible for the backend streams to see the deletion of a row from the old table but not its insertion into the new table, even though the insertion happens first, and the row would be missed. \psharp managed to discover this bug in a matter of seconds. The MigratingTable developers spent just 10 minutes analyzing the trace to diagnose what was happening; although, this was after days of experience analyzing traces and after they had extended the logging of \psharp with additional trace information specific to MigratingTable. \psharp only outputs trace information related to its communicating state machines, but the trace information is easy to extend and was done in all our case studies.

### 2016-01-01 migrating.tex
Edit MigratingTable sections.

Live Table Migration (MigratingTable) is a library designed to \emph{transparently} migrate a data set between tables in an Azure Storage service \emph{while} an application is accessing this data set. This case study is particularly interesting because MigratingTable was \emph{co-developed} with its \psharp test harness. This co-development process resulted in the discovery of several tricky bugs before the system went into production (see \S\ref{sec:eval}). The MigratingTable testing effort differs from the vNext case study: the vNext developers focused on a single liveness property, while the MigratingTable test harness checks complete compliance with an interface (safety) specification.

MigratingTable provides a \emph{virtual table} that has an interface similar to an ordinary Azure table. This interface is named \texttt{IChainTable}. A virtual table is backed by a pair of \emph{old} and \emph{new} \emph{backend} tables, both of which implement \texttt{IChainTable}. A \emph{migrator} job is responsible for moving all data from the old table to the new table in the background. Meanwhile, each read and write operation issued to the virtual table is translated to a sequence of reads and writes on the backend tables according to a protocol, which intends to guarantee that when multiple applications issue \emph{input} operations to virtual tables that are backed by the same backend tables, the output is compliant with the specification of \texttt{IChainTable} for the combined \emph{input history} as if all the operations were done on a single table. The goal of using \psharp was to systematically test this property.

@@

There are two main challenges behind testing MigratingTable: (i) the system is highly concurrent; and (ii) there are many possible input histories. The developers could have chosen specific input histories to test, but they were not confident that this approach would be effective in catching bugs, especially because concurrency increases the potential for difficult-to-foresee interactions between seemingly independent parts of the code. Instead, the developers created a reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, which is backed by a single in-memory table. After executing an arbitrary input history on both implementations, the outputs can be compared to ensure that MigratingTable produces the same output as \texttt{SpecTable}.

Figure~\ref{fig:mockedmigration} illustrates the \psharp test harness of MigratingTable. The harness consists of a set of \texttt{Service} machines that contain identically configured instances of MigratingTable; a \texttt{Tables} machine, which contains the old and new tables, and an instance of \texttt{SpecTable}; and a \texttt{Migrator} machine that performs the background migration. Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which then sends these operations to the \texttt{Tables} machine.

The developers instrumented MigratingTable to report the intended \emph{linearization point} of each input operation on its corresponding virtual table. This allows the \psharp test harness to keep \texttt{SpecTable} in sync with the virtual table: the \texttt{Service} machine sends each input operation to \texttt{SpecTable} when the linearization point on the virtual table is reached. The output of the operation on the virtual table is then checked against its output on \texttt{SpecTable}.

Conceptually, the \psharp test harness randomly chooses (uniformly) an input history from a set that contains all possible input histories under a certain bound \PTComment{What bound?}. Assuming that any given bug in MigratingTable can manifest under at least one of the input histories in the set (the small scope hypothesis), then an execution of the test harness must have a non-zero probability of detecting the bug.

\psharp captures and controls all the interleaving in the system, and systematically explores them. To ensure that \texttt{SpecTable} is never observed to be out of sync with the backend tables, which could lead to false positives, after the \texttt{Tables} machine processes a backend operation, it enters a state that defers further backend operations until MigratingTable has reported whether the backend operation was a linearization point and (if so) the operation on \texttt{SpecTable} has been completed.

>>>>

% Remove the \emph on ``transparently'' and ``while'' again.  Pantazis, I'm
% imagining you were trying to emphasize what's interesting about the problem,
% but I think suggesting that this idea is new to the reader is insulting.
% ~ Matt 2015-12-31
Live Table Migration (MigratingTable) is a library designed to transparently migrate a data set between tables in the Azure storage service while an application is accessing this data set. This case study is particularly interesting because MigratingTable was \emph{co-developed} with its \psharp test harness. This co-development process resulted in the discovery of several tricky bugs before the system went into production (see \S\ref{sec:eval}). The MigratingTable testing effort differs from the vNext case study: the vNext developers focused on a single liveness property, while the MigratingTable test harness checks complete compliance with an interface (safety) specification.

MigratingTable provides a \emph{virtual table} that has an interface
similar to an ordinary Azure table. This interface is
named \texttt{IChainTable}. A virtual table is backed by a pair
of \emph{old} and \emph{new} \emph{backend} tables, both of which
implement \texttt{IChainTable}. A \emph{migrator} job is responsible
for moving all data from the old table to the new table in the
background. Meanwhile, each \emph{input} read and write operation
issued to the virtual table is implemented via a sequence
of reads and writes on the backend tables according to a protocol,
which intends to guarantee that when multiple applications issue input
operations to different MigratingTable instances with the same backend
tables, the output of these operations complies with the specification
of \texttt{IChainTable} for the combined \emph{input history} as if
all the operations were done on a single table. The goal of
using \psharp was to systematically test this property.

@@

There are two main challenges behind testing MigratingTable: (i) the system is highly concurrent; and (ii) the input operations accept many parameters that affect the behavior in different ways. The developers could have chosen specific input histories to test, but they were not confident that they would cover the combinations of parameters that might trigger bugs. Instead, they used the \psharp API to sample from a distribution over input histories that they defined, which allowed all of the relevant parameters to vary within certain limits. They created a reference implementation of the \texttt{IChainTable} specification, called \texttt{SpecTable}, and compared the output of MigratingTable on the chosen input history to that of a \texttt{SpecTable} instance called the \emph{reference table}. The \texttt{SpecTable} implementation was reused for the backend tables of MigratingTable.
% I'm removing the claim that each bug is caught with positive probability per
% run because it's too awkward to explain here and it's nothing specific to
% MigratingTable.  Also, we can't quite appeal to the small scope hypothesis in
% its original form because I hand-picked which parameters to vary and what
% limits to use.  The value of the input generator should be clear enough from
% the above. ~ Matt 2015-12-31

Figure~\ref{fig:mockedmigration} illustrates the \psharp test harness of MigratingTable. The harness consists of a \texttt{Tables} machine that contains the backend tables and the reference table and serializes all operations on these tables; a set of \texttt{Service} machines that contain identically configured instances of MigratingTable; and a \texttt{Migrator} machine that performs the background migration. Each \texttt{Service} machine issues a random sequence of input operations to its MigratingTable, which sends the necessary operations to the backend tables via \psharp events. The developers instrumented MigratingTable to report the \emph{linearization point} of each input operation, i.e., the time at which it takes effect on the single virtual table, so the test harness could perform the operation on the reference table at the same time. (More precisely, after processing each backend operation, the \texttt{Tables} machine enters a \psharp state that blocks all further work until MigratingTable reports whether the backend operation was the linearization point and, if so, the input operation has been completed on the reference table. This way, the rest of the system never observes the reference table to be out of sync with the virtual table.)
% Note: If the tables were observed to be out of sync, it could cause both false
% positives and false negatives. ~ Matt 2015-12-31

For further information about MigratingTable and its \psharp test harness, we refer the reader to the source repository~\cite{migratingtable-src}.
